---
title: "Marker Genes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressMessages({
  library(tidySingleCellExperiment)
  library(ggplot2)
  library(hrbrthemes)
  library(openxlsx)
  library(scater)
  library(scran)
  library(pheatmap)
  library(ggplotify)
  library(viridis)
})


# load utils functions
source("../code/general_purpose_code.R")
```

# Reload the data

```{r Reload the data}
sce <- readRDS(file="../data/EOS_Files/EDI_EOS5_sce.rds")

colLabels(sce) <- sce$kgraph_clusters_walktrap_harmony
```

# UMAP coloured by clusters and protocol {.tabset}

```{r Visualisation of clusters, fig.width=16, fig.height=10, results="asis"}
cat("\n\n## UMAP coloured by cluster \n\n")
plotReducedDim(sce, "UMAP", colour_by="label", text_by="label", point_size=0.5) +
  labs(title="UMAP plot coloured by clusters") +
  theme_ipsum_rc()

cat("\n\n## UMAP coloured by Protocol \n\n")
plotReducedDim(sce, "UMAP", colour_by="Protocol", point_size=0.5) +
  labs(title="UMAP plot coloured by protocol") + 
  theme_ipsum_rc() +
  get_protocol_colours(levels(sce$Protocol))
```

```{r prepare sheet to save data to}
wb <- createWorkbook()
addWorksheet(wb, "Intro")
writeData(wb, "Intro", data.frame("Marker genes obtained using only the highly-variable genes"), colNames=FALSE)
```


# Marker Genes for unsupervised clusters

Marker genes can help us identify clusters as specific cell types and also discover smaller changes between subclusters (e.g. changes in activation or differentiation state)


## Marker genes for each cluster using the t-test {.tabset}

```{r Marker Genes upregulated t-test, fig.height=10, fig.width=7, results="asis"}
sce_marker_genes <- sce[rowData(sce)$is_hvg,]
rownames(sce_marker_genes) <- rowData(sce_marker_genes)$Symbol

markers.all.up <- findMarkers(sce_marker_genes, 
                              test.type="t",
                              direction="up", 
                              block=sce_marker_genes$Sample)

for (i in 1:length(unique(colLabels(sce_marker_genes)))){
  
  cat(paste0("### Cluster ", i, "\n"))
  
  interesting.all.up <- markers.all.up[[i]]
  interesting.all.up <- interesting.all.up[interesting.all.up$Top <=10,]
  logFCs <- getMarkerEffects(interesting.all.up, prefix="logFC")
  logFCs[is.na(logFCs)] <- 0 # set all NAs to 0
  
  # write to excel
  sheet_name <- paste0("ttest_pairs_cluster_", i)
  addWorksheet(wb, sheet_name)
  writeData(wb, sheet_name, logFCs, rowNames=TRUE)
  
  pheatmap(logFCs, breaks=seq(-3, 3, length.out=101),
           main=paste0("Cluster ",i, ": Marker genes, corrected for batch effects \n (upregulated only)")) + 
    theme_ipsum_rc()
  
  cat("\n\n")
}
```

## Heatmap for t-test results

```{r pheatmap for t-test results, fig.width=16, fig.height=40}
top_genes_ttest <- lapply(markers.all.up, function(marker_set){
  marker_set %>%
    as.data.frame() %>%
    filter(Top <= 5) %>%
    rownames()
  }) %>% 
  unlist(use.names=F) %>% 
  unique()

# Matrix to plot the heatmap of expression values. 
# To avoid a too big plot we only keep every 10th cell
sce_plot <- sce_marker_genes[top_genes_ttest, seq(1, dim(sce_marker_genes)[2], 20)]
# We sort by cluster
sce_plot <- sce_plot[,order(sce_plot$label)]
# Anotations
column_annotations <- data.frame("Cluster"=factor(sce_plot$label), row.names=colnames(sce_plot))
# clip heatmap values to ensure that colour contrast is high
sce_plot <- assay(sce_plot, "logcounts")
sce_plot[sce_plot >=3] = 3

as.ggplot(pheatmap(sce_plot, annotation_col=column_annotations,
         show_rownames=TRUE, show_colnames=FALSE, cluster_cols=FALSE, silent=TRUE,
         color=inferno(50))) +
  labs(title="Gene signatures per cluster identified via t-test",
       subtitle="Top 5 differentially expressed genes per cluster; values are log-counts clipped at 3") +
  theme_ipsum_rc()+ 
  theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank())
```


## Marker genes for each cluster using the Wilcox {.tabset}

The Wilcoxon test makes pairwise comparisons between to groups (clusters) which is proportional to the area und the curve (AUC). The scores represent the likelihood of a random cell of another cluster having a higher expression than the reference cluster. Scores of 0 and 1 indicate that the distributions are completely separated.

```{r marker genes wilcox upregulated, fig.height=10, fig.width=7, results="asis"}
markers.all.up <- findMarkers(sce_marker_genes, 
                              test.type="wilcox",
                              pval.type="any",
                              direction="up", 
                              block=sce_marker_genes$Sample)

for (i in 1:length(unique(colLabels(sce_marker_genes)))){
  
  cat(paste0("### Cluster ", i, "\n"))
  
  interesting.all.up <- markers.all.up[[i]]
  interesting.all.up <- interesting.all.up[interesting.all.up$Top <=10,]
  logFCs <- getMarkerEffects(interesting.all.up, prefix="AUC")
  logFCs[is.na(logFCs)] <- 0 # set all NAs to 0
  
   # write to excel
  sheet_name <- paste0("wilcox_pairs_cluster_", i)
  addWorksheet(wb, sheet_name)
  writeData(wb, sheet_name, logFCs, rowNames=TRUE)
  
  pheatmap(logFCs, breaks=seq(0, 1, length.out=101),
           main=paste0("Cluster ",i, ": Marker genes, corrected for batch effects \n (upregulated only)")) + 
    theme_ipsum_rc()
  
  cat("\n\n")
}
```

## Heatmap for Wilcoxon test results

```{r pheatmap for wilcox results, fig.width=16, fig.height=40}
top_genes_wilcox <- lapply(markers.all.up, function(marker_set){
  marker_set %>%
    as.data.frame() %>%
    filter(Top <= 5) %>%
    rownames()
}) %>% 
  unlist(use.names=F) %>% 
  unique()

# Matrix to plot the heatmap of expression values. 
# To avoid a too big plot we only keep every 10th cell
sce_plot <- sce_marker_genes[top_genes_wilcox, seq(1, dim(sce_marker_genes)[2], 20)]
# We sort by cluster
sce_plot <- sce_plot[,order(sce_plot$label)]
# Anotations
column_annotations <- data.frame("Cluster"=factor(sce_plot$label), row.names=colnames(sce_plot))
# clip heatmap values to ensure that colour contrast is high
sce_plot <- assay(sce_plot, "logcounts")
sce_plot[sce_plot >=3] = 3

as.ggplot(pheatmap(sce_plot, annotation_col=column_annotations,
         show_rownames=TRUE, show_colnames=FALSE, cluster_cols=FALSE, silent=TRUE,
         color=inferno(50))) +
  labs(title="Gene signatures per cluster identified via wilcox test",
       subtitle="Top 5 differentially expressed genes per cluster; values are log-counts clipped at 3") +
  theme_ipsum_rc()+ 
  theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank())
```


```{r save marker genes to excel}
filename_marker_results <- "../docs/output/EDI_marker_results_hvg_genes.xlsx"
saveWorkbook(wb, file = filename_marker_results, overwrite = TRUE)
```