---
title: "Cell Cycle Analysis"
author: "Dominique Paul"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressMessages({
  library(RCurl)
  library(Seurat)
  library(hrbrthemes)
  library(ggplot2)
  library(SingleCellExperiment)
  library(dplyr)
  library(scater)
})

theme_set(theme_ipsum_rc())

# this script contains the function 'plot_stratified_views' which makes plotting 
# dimension reductions easier and is used throughout the project
source("../code/general_purpose_code.R")
```


# Load data

```{r load-data}
sce <- readRDS(file="../data/EOS_Files/EDI_EOS3_sce.rds")
rownames(sce) <- rowData(sce)$ID # might need this for the analysis
```

We download the marker genes for the cell cycle stages from a github repository. The data comes in a data frame. We split the genes into lists according to the cell cycle phase.

```{r get cell cycle genes}
cc_file <- getURL("https://raw.githubusercontent.com/hbc/tinyatlas/master/cell_cycle/Homo_sapiens.csv") 
cell_cycle_genes <- read.csv(text = cc_file)

head(cell_cycle_genes)

s_genes <- cell_cycle_genes %>%
  dplyr::filter(phase == "S") %>%
  pull("geneID")

g2m_genes <- cell_cycle_genes %>%
  dplyr::filter(phase == "G2/M") %>%
  pull("geneID")
```


# Identify cell cycle phases

The function we are using for the identification of the cell cycle is a Seurat function. Seurat is a data format similar to the Single Cell Experiment object we generally use in the other scrips of this analysis. So we convert our Single Cell Experiment object into a Seurat object and run the identification of the cell cycle phase. We save the results to our Single Cell Experiment object as we will be using this object class it in our remaining analysis.


```{r cell cycle scoring with seurat, warning=FALSE}
sce_seurat <- as.Seurat(sce, counts="counts", data="logcounts")

sce_seurat <- CellCycleScoring(sce_seurat,
                               g2m.features = g2m_genes,
                               s.features = s_genes)

sce$phase <- sce_seurat$Phase
```

# Visualisation

## Coloured by clusters label {.tabset}

We show the plots once to inspect the clusters. It could be interesting to see that certain clusters might have been formed based on the similarity of the cell cycle phase.

```{r plot UMAP, results="asis", fig.width=16, fig.height=10}
cat("\n\n### UMAP \n\n")
plotReducedDim(sce, "UMAP", colour_by="kgraph_clusters_harmony", text_by="kgraph_clusters_harmony", point_size=0.5) +
  theme_ipsum_rc()
```

## Coloured by cell cycle phase {.tabset}

```{r visualise-phases-umap by protocol, fig.width=16, fig.height=10, results="asis"}
plot_stratified_views(sce, "UMAP", stratify_by="Protocol", colour_by="phase", text_by="kgraph_clusters_harmony", tabset="###", point_size=0.5)
```

## Coloured by protocol & grouped by cell cycle phase {.tabset}

```{r visualise-phases-umap by phase, fig.width=16, fig.height=10, results="asis"}
plot_stratified_views(sce, "UMAP", stratify_by="phase", colour_by="Protocol", text_by="kgraph_clusters_harmony", tabset="###", point_size=0.5)
```


# Cluster-phase composition (stacked barplot) {.tabset}

```{r average doublet score 1, results="asis", warning=FALSE, fig.width=12, fig.height=10}
# Cluster_50
cat("\n\n## By Cluster \n")
as.data.frame(colData(sce)) %>%
  group_by(kgraph_clusters_harmony, phase) %>%
  summarise(Freq=n()) %>% 
  ggplot(aes(x=kgraph_clusters_harmony, y=Freq, fill=phase, label=Freq)) +
         geom_bar(stat="identity", position="fill") +
  labs(title="Frequency of cell phase by cluster",
       subtitle="Labels represent absolute number of cells per group",
       x="Cluster",
       y="Frequency") + 
  geom_text(size=3, position = position_fill(vjust=0.5))

# Protocol
cat("\n\n## By Protocol \n")
as.data.frame(colData(sce)) %>%
  group_by(Protocol, phase) %>%
  summarise(Freq=n()) %>%
  ggplot(aes(x=Protocol, y=Freq, fill=phase, label=Freq)) +
         geom_bar(stat="identity", position="fill") +
  labs(title="Frequency of cell phase by Protocol",
       subtitle="Labels represent absolute number of cells per group",
       x="Cluster",
       y="Frequency") + 
  geom_text(size=3, position = position_fill(vjust=0.5))


# Sample
cat("\n\n## By Sample \n")
as.data.frame(colData(sce)) %>%
  group_by(Sample, phase) %>%
  summarise(Freq=n()) %>%
  ggplot(aes(x=Sample, y=Freq, fill=phase, label=Freq)) +
         geom_bar(stat="identity", position="fill") +
  labs(title="Frequency of cell phase by Sample",
       subtitle="Labels represent absolute number of cells per group",
       x="Cluster",
       y="Frequency") + 
  geom_text(size=3, position = position_fill(vjust=0.5)) +
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())
```


# Cluster-phase composition (counts) {.tabset}

```{r average doublet score 2, results="asis", warning=FALSE, fig.width=12, fig.height=10}
# Cluster_50
cat("\n\n## By Cluster \n")
as.data.frame(colData(sce)) %>%
  group_by(kgraph_clusters_harmony, phase) %>%
  summarise(Freq=n()) %>% 
  ggplot(aes(x=kgraph_clusters_harmony, y=Freq, fill=phase, label=Freq)) +
         geom_bar(stat="identity") +
  labs(title="Frequency of cell phase by cluster",
       subtitle="Labels represent absolute number of cells per group",
       x="Cluster",
       y="Frequency") + 
  geom_text(size=3)

# Protocol
cat("\n\n## By Protocol \n")
as.data.frame(colData(sce)) %>%
  group_by(Protocol, phase) %>%
  summarise(Freq=n()) %>%
  ggplot(aes(x=Protocol, y=Freq, fill=phase, label=Freq)) +
         geom_bar(stat="identity") +
  labs(title="Frequency of cell phase by Protocol",
       subtitle="Labels represent absolute number of cells per group",
       x="Cluster",
       y="Frequency") + 
  geom_text(size=3)


# Sample
cat("\n\n## By Sample \n")
as.data.frame(colData(sce)) %>%
  group_by(Sample, phase) %>%
  summarise(Freq=n()) %>%
  ggplot(aes(x=Sample, y=Freq, fill=phase, label=Freq)) +
         geom_bar(stat="identity") +
  labs(title="Frequency of cell phase by Sample",
       subtitle="Labels represent absolute number of cells per group",
       x="Cluster",
       y="Frequency") + 
  geom_text(size=3) +
  theme(axis.text.x = element_text(angle = 45,hjust=1), axis.ticks.x=element_blank())
```


# Save data 

```{r save data}
rownames(sce) <- rowData(sce)$Symbol

file_path <- "../data/EOS_Files/EDI_EOS4_sce.rds"
saveRDS(sce, file = file_path)
upload_file_to_iSEE(file_path)
```

