---
title: "Normalisation, HVGs, Dimensionality Reduction"
author: "Dominique Paul"
date: "11/1/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressMessages({
  library(tidySingleCellExperiment)
  library(scran)
  library(BiocParallel)
  library(scater)
  library(parallel)
  library(hrbrthemes)
  library(kableExtra)
  library(intrinsicDimension)
  library(Seurat)
})

# Setting for parallel computing
n_workers <- min(detectCores()-1, 16)
RhpcBLASctl::blas_set_num_threads(n_workers)
bpparam <- BiocParallel::MulticoreParam(workers=n_workers, RNGseed = 123)

# load utils functions
source("../code/general_purpose_code.R")
```


# Load data

```{r}
sce <- readRDS(file = "../data/EOS_Files/EDI_EOS1_sce.rds")
```

# Normalisation via SC Transform 

```{r sctransform, fig.width=12, fig.height=8, warning=FALSE}
sce <- logNormCounts(sce)

sce_seurat <- as.Seurat(sce, counts="counts", data="logcounts")

seurat_normalized = sce_seurat %>%
  Seurat::SCTransform(verbose = FALSE, assay="originalexp", min_cells=1)


sct_HVGs <- seurat_normalized@assays$SCT@var.features

assay(sce, "corrected") <- seurat_normalized@assays$SCT@counts
assay(sce, "logcounts") <- seurat_normalized@assays$SCT@data
rowData(sce)[, "sct_HVG"] <- FALSE
rowData(sce)[sct_HVGs, "sct_HVG"] <- TRUE


# Run PCA
set.seed(100)
sce <- runPCA(sce, subset_row=rowData(sce)$sct_HVG)

plotReducedDim(sce, "PCA", colour_by=I(sce$log_total), point_size=0.1) +
  labs(title="Correlation between library size and principal components",
       subtitle="Coloured by library size factors with colouring capped at 5") +
  scale_colour_gradientn(#limits = c(0,5),  
                         colours=c("navyblue", "darkmagenta", "darkorange1"),) +
  theme_ipsum_rc()

# Check for correlation of PCA with log_total
cor.test(reducedDim(sce, "PCA")[,"PC1"], sce$log_total)
cor.test(reducedDim(sce, "PCA")[,"PC2"], sce$log_total)
```



# Identification of highly variable genes

Gene selection is important for the downstream analysis, such as clustering and dimensionality reduction. A too high number of genes can slow down computation speed and by including genes which aren't interesting from a biological perspective we, more importantly, add more noise to the analysis. In general, there is a trade-off between retaining biological signals and increasing the noise form irrelevant genes.

We model the mean and the variance for each gene. Subsequently a trend is fitted to the data. Based on this trend we identify which genes strongly deviate from the trend and calculate the statistical significance: the p-value transformed into an false discovery rate (FDR) of the given result. We mark all genes with an FDR value 0.05 as a highly variable gene.


## Show highly variable genes

```{r gene variation with adjusted data , fig.width=14, fig.height=10}
dec.sce <- modelGeneVar(sce, block=sce$Sample)
#fit.vs <- metadata(dec.sce)

hvg_sub <- getTopHVGs(dec.sce, fdr.threshold=0.05)

mean_var_comb_sub <- purrr::map(unique(sce$Sample), ~ {
 dec.sce$per.block[[.x]] %>% 
  as_tibble %>% 
  mutate(row_names = rownames(dec.sce$per.block[[.x]]),
         is_hvg = row_names %in% hvg_sub,
         Sample = .x)
  }) %>% 
  purrr::reduce(rbind)

mean_var_comb_sub %>% 
  ggplot() +
    geom_point(aes(x = mean, y = total, color= is_hvg)) +
    geom_line(aes(x=mean, y= tech)) +
    labs(y="Variance",x="Mean expression") +
    facet_wrap(~Sample) +
    scale_color_manual(values=c("orange","black")) +
    labs(x="Mean of log-expression", 
         y="Variance of log-expression",
         title="Variance of the expression values as a function of the mean")

rowData(sce)$is_hvg <- FALSE
rowData(sce)[hvg_sub,]$is_hvg <- TRUE
```

# {-}

## Tabular overview of highly variable genes

```{r variance}
# Ordering by most interesting genes for inspection.
dec.sce[order(dec.sce$bio, decreasing=TRUE),] %>%
       as.data.frame() %>%
       dplyr::select(c("mean", "total", "tech", "bio", "p.value", "FDR")) %>%
  head(n=10) %>%
  kable(digits=5) %>%
  kable_styling(full_width=TRUE)
```


# Dimension reduction

We create a principal components dimensionality reduction of our data. We then use `maxLikGlobalDimEst` from the `intrinsicDimension` package to choose a number of dimensions that captures a big part of the variety in the data and at the same reduces the complexity of further operations. We use these reduced dimensions to calculate a UMAP (Uniform Manifold Approximation) dimensionality reduction.

However, in the next script we run the batch integration which we can then use for a much better UMAP.

```{r dimension reduction}
# run PCA
set.seed(100)
sce_tmp <- runPCA(sce, name = "PCA", subset_row=rowData(sce)$is_hvg)

# subset dimensions for reduced representation
ndims <- intrinsicDimension::maxLikGlobalDimEst(as.matrix(reducedDim(sce_tmp, "PCA")), k=20)
reducedDim(sce_tmp,"PCA_reduced") <- reducedDim(sce_tmp,"PCA")[,seq_len(ceiling(ndims$dim.est))]

# run UMAP
set.seed(100)
sce_tmp <- runUMAP(sce_tmp, name = "UMAP", dimred = "PCA_reduced", subset_row=rowData(sce_tmp)$is_hvg, n_neighbors=50, min_dist = 0.1)


# add all reduced dimensions to the original sce object
reducedDim(sce, "PCA") <- reducedDim(sce_tmp, "PCA")
reducedDim(sce, "PCA_reduced") <- reducedDim(sce_tmp, "PCA")[,seq_len(ceiling(ndims$dim.est))]
reducedDim(sce, "UMAP") <- reducedDim(sce_tmp,"UMAP")

cor.test(reducedDim(sce, "PCA")[,"PC1"], sce$log_total)
```


# Plot PCA and UMAP (no batch correction yet)

The data is not yet batch-corrected, so we don't expect any interesting biological differences to be reflected in the first reduced dimension plots.

## PCA {.tabset}

```{r PCA plot, fig.width=14, fig.height=10, results="asis"}
# by Sample
cat("\n\n### PCA by sample \n")
set.seed(100)
shuffle <- sample(seq_len(dim(sce)[2]))
plotReducedDim(sce[,shuffle], "PCA", colour_by = "Sample") +
  labs(title="PCA coloured by sample") +
  theme_ipsum_rc() +
  get_sample_colours(levels(sce$Sample))

# by protocol
cat("\n\n### PCA by protocol \n")
set.seed(100)
shuffle <- sample(seq_len(dim(sce)[2]))
plotReducedDim(sce[,shuffle], "PCA", colour_by = "Protocol") +
  labs(title="PCA coloured by protocol") +
  theme_ipsum_rc() +
  get_protocol_colours(levels(sce$Protocol))
```

## UMAP {.tabset}

We see that the UMAP clearly does not filter for biological features like cell types before applying any batch correction.

```{r UMAP, fig.width=14, fig.height=10, results="asis"}
# by sample
cat("\n\n### UMAP by sample \n")
set.seed(100)
shuffle <- sample(seq_len(dim(sce)[2]))
plotReducedDim(sce[,shuffle], "UMAP", colour_by = "Sample", point_size=0.5) +
  labs(color="Sample",
       title="UMAP (uncorrected) coloured by Sample") +
  theme_ipsum_rc() +
  get_sample_colours(levels(sce$Sample))

# by protocol
cat("\n\n### UMAP by protocol \n")
set.seed(100)
shuffle <- sample(seq_len(dim(sce)[2]))
plotReducedDim(sce[,shuffle], "UMAP", colour_by = "Protocol", point_size=0.5) +
  labs(title="UMAP (uncorrected) coloured by Protocol", color="Protocol") +
  theme_ipsum_rc()
  get_protocol_colours(levels(sce$Protocol))
```


# Save data and upload to iSEE

```{r save and upload}
file_path <- "../data/EOS_Files/EDI_EOS2_sce.rds"
saveRDS(sce, file = file_path)
upload_file_to_iSEE(file_path)
```




