---
title: "Differential Abundance"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressMessages({
  library(tidySingleCellExperiment)
  library(edgeR)
  library(ggplot2)
  library(hrbrthemes)
  library(scater)
})

# load utils functions
source("../code/general_purpose_code.R")
```

# Load data

```{r load-data, fig.width=16, fig.height=10}
sce <- readRDS(file="../data/EOS_Files/EDI_EOS5_sce.rds")
```


# Differential Abundance of cell types (using own clusters)

```{r code setup}
test_col <- "Protocol"
plotvars <- c("Protocol")
label <- "manual_labels_coarse"


# Cross-tabulate the label and the samples
abundances <- table(colData(sce)[[label]], colData(sce)[["Sample"]])
# returns a copy of the table with its table class attribute removed
abundances <- unclass(abundances) 

# We attach the samples and protocol columns as additiona metadata
extra.info <- colData(sce)[match(colnames(abundances), sce$Sample),c("Sample", plotvars)]
extra.info[[test_col]] <- stringr::str_replace_all(extra.info[[test_col]]," ","_")

# setup
y.ab <- DGEList(abundances, samples=extra.info, group=factor(extra.info[[test_col]]))

# filter out cell types which don't have enough cell counts
keep <- filterByExpr(y.ab, group=factor(y.ab$samples[[test_col]]))
y.ab <- y.ab[keep,]
summary(keep)
```



```{r design matrix and basic plots}
design_matrix <- rlang::eval_tidy(rlang::expr(
  model.matrix(!!as.formula(
    paste("~",paste(c("0",paste0("factor(",test_col,")")),
                    collapse = " + "))),
    y.ab$samples)
  ))

# the columns of the design matrix contain 'factor(Protocol)' in front 
# of the protocol name. We remove this
colnames(design_matrix) <- stringr::str_extract(colnames(design_matrix),"(?<=\\))([:graph:]| )+|P_[:graph:]+")
print(design_matrix)

# Create a contrast matrix out of the design matrix
combs_contr <- combn(sort(unique(y.ab$samples[[test_col]])), 2)
contrast_names <- apply(combs_contr,2,function(x) paste(x,collapse = "-"))
contrast <- makeContrasts(contrasts=contrast_names, 
                           levels = sort(c(unique(y.ab$samples[[test_col]]))))
print(contrast)

# estimate dispersion
# Plot the biological coefficient of variation (BCV) of a cell type against its abundance (in log2 counts per million).
y.ab <- estimateDisp(y.ab, design_matrix, trend="none")
plotBCV(y.ab, cex=1)

# fit
# Fit a quasi-likelihood negative binomial generalized log-linear model to count data. 
# Conduct cell type specific statistical tests for a given coefficient or contrast.
fit.ab <- glmQLFit(y.ab, design_matrix, robust=TRUE, abundance.trend=FALSE)
# Plot the per cell type quasi-likelihood dispersion against the cell type abundance (in log2 counts per million).
plotQLDisp(fit.ab, cex=1)

da_objects <- list()
```

# Individual tests {.tabset}

We test the fresh protocol versus other protocols for statistical changes in cell type abundances. The first table indicates how many cell types were identified to exhibit a statistically significant change. The second table shows the exact log fold changes and the associated p-value and false-discovery rate.

## Test Fresh vs. Tabib

```{r Fresh vs. Tabib}
da_objects[["Fresh-Tabib_et_al"]] <- res_fresh_lafa <- glmQLFTest(fit.ab, contrast = contrast[,"Fresh-Tabib_et_al", drop=FALSE])
knitr::kable(summary(decideTests(res_fresh_lafa)))

as.data.frame(topTags(res_fresh_lafa, n=20)[,c("logFC", "logCPM",	"F", "PValue", "FDR")]) %>%
  knitr::kable(digits=2)
```

## Test Fresh vs. He et al.

```{r Fresh vs. He}
da_objects[["Fresh-He_et_al"]] <- res_fresh_he_et_al <- glmQLFTest(fit.ab, contrast = contrast[,"Fresh-He_et_al", drop=FALSE])
knitr::kable(summary(decideTests(res_fresh_he_et_al)))

as.data.frame(topTags(res_fresh_he_et_al, n=20)[,c("logFC", "logCPM",	"F", "PValue", "FDR")]) %>%
  knitr::kable(digits=2)
```

## Test Fresh vs. Sole-Boldo

```{r Fresh vs. Sole-Boldo}
da_objects[["Fresh-Sole_Boldo_et_al"]] <- res_fresh_sole_boldo <- glmQLFTest(fit.ab, contrast = contrast[,"Fresh-Sole_Boldo_et_al", drop=FALSE])
knitr::kable(summary(decideTests(res_fresh_sole_boldo)))

as.data.frame(topTags(res_fresh_sole_boldo, n=20)[,c("logFC", "logCPM",	"F", "PValue", "FDR")]) %>%
  knitr::kable(digits=2)
```

## Test He vs. Sole-Boldo

```{r He vs. Sole-Boldo}
da_objects[["He_et_al-Sole_Boldo_et_al"]] <- res_he_sb <- glmQLFTest(fit.ab, contrast = contrast[,"He_et_al-Sole_Boldo_et_al", drop=FALSE])
knitr::kable(summary(decideTests(res_he_sb)))

as.data.frame(topTags(res_he_sb, n=20)[,c("logFC", "logCPM",	"F", "PValue", "FDR")]) %>%
  knitr::kable(digits=2)
```

## Test He vs. Tabib

```{r He vs. Tabib}
da_objects[["He_et_al-Tabib_et_al"]] <- res_he_tabib <- glmQLFTest(fit.ab, contrast = contrast[,"He_et_al-Tabib_et_al", drop=FALSE])
knitr::kable(summary(decideTests(res_he_tabib)))

as.data.frame(topTags(res_he_tabib, n=20)[,c("logFC", "logCPM",	"F", "PValue", "FDR")]) %>%
  knitr::kable(digits=2)
```

## Test Sole-Boldo vs. Tabib

```{r He vs. Tabib}
da_objects[["Sole_Boldo_et_al-Tabib_et_al"]] <- res_sb_tabib <- glmQLFTest(fit.ab, contrast = contrast[,"Sole_Boldo_et_al-Tabib_et_al", drop=FALSE])
knitr::kable(summary(decideTests(res_sb_tabib)))

as.data.frame(topTags(res_sb_tabib, n=20)[,c("logFC", "logCPM",	"F", "PValue", "FDR")]) %>%
  knitr::kable(digits=2)
```


## Save the tables

```{r}
folder_path <- "../output/review_figures/DA/"

for (comparison in names(da_objects)){
  df <- as.data.frame(topTags(da_objects[[comparison]], n=20)[,c("logFC", "logCPM", "PValue", "FDR")])
  # Save as image
  kable(df, digits=4) %>%
    kable_styling() %>%
    save_kable(file = paste0(folder_path, "DA_", comparison, ".jpg"))
}


# Write as c

```

# Test all contrasts

For how many cell types is there a significant change in abundance?

```{r all contrasts}
# Fit a quasi-likelihood negative binomial generalized log-linear model to count data. Conduct genewise statistical tests for a given coefficient or contrast.
res <- glmQLFTest(fit.ab, contrast = contrast)
knitr::kable(summary(decideTests(res)))

knitr::kable(topTags(res, n=20), digits=2)
```

```{r plot abundances, fig.width=16, fig.height=10}
# plot abundances
purrr::map(rownames(topTags(res, n=12)), ~ {
  tibble::tibble(counts = y.ab$counts[.x,],
                 abundances = counts/y.ab$samples$lib.size,
                 group = y.ab$samples$group,
                 Cluster = .x,
                 Protocol = y.ab$samples$Protocol)}) %>% 
  purrr::reduce(rbind) %>% 
  dplyr::mutate(Cluster = factor(Cluster, levels = rownames(topTags(res, n=12)))) %>% 
  ggplot() + 
  geom_boxplot(aes(y=abundances,x=Protocol, color=Protocol)) +
  geom_point(aes(y=abundances,x=Protocol, color=Protocol), size = 3) +
  # scale_color_manual(values=sample_cols(unique(y.ab$samples$Protocol),n_split = 4, palette=viridis::viridis)) +
  theme(axis.text.x = element_text(angle=45, hjust = 1), strip.text.x = element_text(family = "Helvetica",face = "bold")) +
  labs(y="relative Abundance", x="") +
  facet_wrap(~Cluster, scales = "free_y",ncol = 4, nrow = 3) +
  get_protocol_colours(levels(sce$Protocol))
```


```{r plot abundances v2, fig.width=20, fig.height=10}
count_table <- table(sce$manual_labels_coarse, sce$Sample)
lib_sizes <- colSums(count_table)

sample_protocol_mapping <- as.data.frame(colData(sce)) %>%
  group_by(Protocol, Sample) %>%
  summarise() %>%
  dplyr::ungroup() %>%
  tibble::column_to_rownames('Sample')



fig11d <- purrr::map(unique(sce$manual_labels_coarse), ~ {
  tibble::tibble(counts = count_table[.x,],
                 abundances = counts/lib_sizes,
                 Protocol= sample_protocol_mapping[levels(sce$Sample), "Protocol"],
                 Cluster = .x)}) %>% 
  purrr::reduce(rbind) %>% 
  dplyr::arrange("Cluster") %>%
  ggplot() + 
  geom_boxplot(aes(y=abundances,x=Protocol, color=Protocol)) +
  geom_point(aes(y=abundances,x=Protocol, color=Protocol), size = 3) +
  theme_ipsum_rc() +
  theme(axis.text.x = element_text(angle=45, hjust = 1), strip.text.x = element_text(family = "Helvetica",face = "bold"),
        legend.position="none") +
  scale_y_continuous(labels = scales::percent) +
  labs(title="Relative cell type abundance", x="", y="") +
  facet_wrap(~Cluster, scales = "free_y",ncol = 5, nrow = 2) +
  get_protocol_colours(levels(sce$Protocol))
fig11d
```

# UMAP proportions

```{r, fig.width=16, fig.height=16}
ret_vals <- plot_label_barplots(sce, "manual_labels_coarse", level1="Protocol", level2=NA, print=F, add_labels=c(FALSE, TRUE))

fig11b1 <- ret_vals$plt_rel +
  get_cell_colours(sce$manual_labels_coarse, type="fill") +
  theme_ipsum_rc() +
  theme(strip.text.y = element_blank(),
        strip.text.x = element_text(size = 10),
        axis.text.x=element_blank(),
        panel.spacing = unit(5, "mm"),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.5,  # Bottom margin
                             l = 20)) +# Left margin
  guides(fill=guide_legend(title="Cell type")) +
  labs(title="Cell type frequency",
       subtitle="Integrated datasets",
       y="",
       x="")
fig11b1

fig11b2 <- ret_vals$plt_abs +
  get_cell_colours(sce$manual_labels_coarse) +
  scale_y_continuous(n.breaks=3) +
  get_cell_colours(sce$manual_labels_coarse, type="fill") +
  theme_ipsum_rc() +
  theme(strip.text.y = element_blank(),
        strip.text.x = element_text(size = 10),
        axis.text.x=element_blank(),
        panel.spacing = unit(5, "mm"),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.5,  # Bottom margin
                             l = 20)) # Left margin 
fig11b2
```

```{r, fig.width=16, fig.height=16}

column <- "manual_labels_coarse"
aggr_by <- "Protocol"
facet_by <- NA
                                
propdf <- table(colData(sce)[[column]], sce[[aggr_by]]) %>% 
  as.data.frame() %>% 
  dplyr::rename(!!column := Var1,
                !!aggr_by := Var2,
                Abundance = Freq) %>% 
  dplyr::group_by(!!sym(aggr_by)) %>% 
  dplyr::filter(Abundance > 0) %>% 
  dplyr::mutate(Proportion = Abundance/sum(Abundance)) %>%
  mutate(Proportion_rounded=paste0(round(Proportion, 3)*100,"%"))
  

fig11c <- propdf %>% 
  ggplot() +
  geom_bar(aes(x = Protocol, y = Proportion, fill=manual_labels_coarse), stat = "identity") +
  geom_text(aes(x = Protocol, y = Proportion, label = Proportion_rounded)) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45,hjust=1),
        strip.text.y = element_text(angle=0,size=7)) +
  labs(title="Fraction of cells by cell type", x="", y="") +
  theme_ipsum_rc() +
  get_cell_colours(unique(propdf$manual_labels_coarse), type="fill") + 
  facet_grid(manual_labels_coarse ~ ., scales = "free_x", space="free") +
  scale_y_continuous(labels = scales::percent, n.breaks=3) +
  theme(panel.spacing = unit(5, "mm"),
        legend.position="bottom",
        plot.margin = margin(t = 30,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.5,  # Bottom margin
                             l = 20))  +# Left margin
  guides(fill=guide_legend(title="",
                            nrow=3,
                            byrow=TRUE, 
                            override.aes = list(size=4, shape = 15, alpha = 1)))
fig11c
```


```{r visualise UMAPs, fig.width=16, fig.height=10, results="asis"}

cat("\n\n## Cell types \n\n")
fig11a1 <- plotReducedDim(sce, "UMAP", colour_by="manual_labels_coarse", text_by="manual_labels_coarse", point_size=0.25, text_size=4) +
  labs(title="Manual labels for high-level clusters",
       subtitle = "Integrated datasets") +
  theme_ipsum_rc() +
  get_cell_colours(levels(sce$manual_labels_coarse)) +
  # theme(legend.position="bottom") +
  guides(color=guide_legend(title="",
                            # nrow=3,
                            byrow=TRUE, 
                            override.aes = list(size=4, shape = 15, alpha = 1)))
fig11a1


cat("\n\n## Protocol \n\n")
fig11a2 <- plotReducedDim(sce, "UMAP", colour_by="Protocol", point_size=0.25) +
  labs(title="UMAP coloured by protocol",
       subtitle = "Integrated datasets") +
  theme_ipsum_rc() +
  get_protocol_colours(levels(sce$Protocol)) +
  # theme(legend.position="bottom") +
  guides(color=guide_legend(title="",
                            #nrow=1,
                            byrow=TRUE, 
                            override.aes = list(size=4, shape = 15, alpha = 1)))

fig11a2
```



# Figure 11.1

```{r figure 11.1, fig.width=10, fig.height=12, warnings=F}
folder_path <- "/Users/dominiquepaul/xRobinson Group/01_Systemic Sclerosis/output/plots/"

figure11_1 <- ggpubr::ggarrange(
  fig11a1,
  fig11a2,
  ncol = 1, nrow = 2,labels = "AUTO"
)

suppressWarnings(figure11_1)

ggsave(paste0(folder_path, "figure11_1.png"),
       plot=figure11_1,
       width=20*0.9,
       height=26*0.9,
       units="in",
       bg="white")
```




# Figure 11.2

```{r supplementary figure 11.2, fig.width=20, fig.height=24, warnings=F}
figure11_2 <- ggpubr::ggarrange(
  ggpubr::ggarrange(fig11b1,fig11b2, ncol = 2),
  #fig11c
  fig11d,
  heights=c(1.5,1),
  ncol = 1, nrow = 2,labels = "AUTO"
)

suppressWarnings(figure11_2)

ggsave(paste0(folder_path, "figure11_2.png"),
       plot=figure11_2,
       width=20*0.9,
       height=26*0.9,
       units="in",
       bg="white")
```


```{r EDI supplemental, fig.width=16, fig.height=10}
EDI_supplemental <- plotReducedDim(sce, "UMAP", colour_by="kgraph_clusters_walktrap_harmony", text_by="kgraph_clusters_walktrap_harmony", point_size=0.25, text_size=4) +
  labs(title="Unsupervised clustering",
       subtitle = "Integrated datasets") +
  theme_ipsum_rc() +
  # get_cell_colours(levels(sce$manual_labels_coarse)) +
  # theme(legend.position="bottom") +
  guides(color=guide_legend(title="",
                            # nrow=3,
                            byrow=TRUE, 
                            override.aes = list(size=4, shape = 15, alpha = 1)))

EDI_supplemental

ggsave(paste0(folder_path, "EDI_supplemental.png"),
       plot=EDI_supplemental,
       width=16*0.9,
       height=10*0.9,
       units="in",
       bg="white")

```
