---
title: "Analysis of heat shock proteins"
author: "Dominique Paul"
date: "`r Sys.Date()`"
output: html_document
---

```{r, include=FALSE}
suppressPackageStartupMessages({
  library(SingleCellExperiment)
  library(muscat)
  library(ggplot2)
  library(ggplotify)
  library(kableExtra)
  library(dplyr)
})
#update.packages(ask=FALSE)
```

# Heat shock proteins

Heat shock proteins (HSPs) are specific proteins that are made when cells are briefly exposed to temperatures above their normal growth temperature. The synthesis of HSPs is a universal phenomenon, occurring in all plant and animal species studied, including humans.

Some people have demonstrated that during tissue dissociation, digestion results in a stress response. Some proteins are produced by cells in response to exposure to stressful conditions (Heat shock proteins). 

In this analysis we examine whether the genes involved in the heat shock are differentially expressed depending on the protocol used. We look at differential states between the protocols. Comparing by cell type makes more sense as differences in average expression could be distorted by different cell type abundances.

The method used for DE in the pseudobulk analysis has a strong impact on the results. The three main options are:
* EdgeR
* Limma
* DESeq2

The genes for HSPs are commonly also termed HSP[...]. The list of genes examined was acquired from the [HGNC database](https://www.genenames.org/data/genegroup/#!/group/582).

## Notes to facilitate coding

The counts matrix includes the raw counts. The logcounts are normalised using scTransform.

Open questions
* Use prefiltered data??
* For which method does normalisation make a difference?
* Use an intercept/samples in design matrix?

```{r hsp0}
# hsps <- c("HSP90AA1", "HSPA9", "HSPD1", "HSPA1A", "HSPA1B", "HSPA8")
hsp_csv <- read.csv("../Metadata/Heat_shock_proteins.csv")
hsps <- hsp_csv$gene_name
cat(hsps)
```

# Load data

```{r hsp2} 
sce_integrated <- readRDS(file="../data/EOS_Files/EDI_EOS5_sce.rds")
sce_fvc <- readRDS(file="../data/EOS_Files/FvC_EOS3_sce.rds")

# table(sce_integrated$manual_labels_coarse)
# table(sce_fvc$manual_labels_coarse)

# We remove some cell types due to low expression levels
# t(table(sce_integrated$manual_labels_coarse, sce_integrated$Sample))
removed_cell_types <- c("Mast cells","Melanocytes/Schwann cells/Neuronal cells") #, "Sweat gland cells")
sce_integrated <- sce_integrated[,!sce_integrated$manual_labels_coarse  %in% removed_cell_types]

# t(table(sce_fvc$manual_labels_coarse, sce_fvc$Protocol))
removed_cell_types_fvc <- c("Schwann cells", "Dendritic cells", "Macrophages")
sce_fvc <- sce_fvc[,!sce_fvc$manual_labels_coarse  %in% removed_cell_types_fvc]
```

Are all of the genes in our data?

```{r hsp 2}
# all hsp genes appear to be present in the data
hsps %in% rowData(sce_integrated)$Symbol # integrated data
hsps %in% rowData(sce_fvc)$Symbol # fresh v culture

### Optionally drop genes from list
# hsps <- hsps[hsps %in% rowData(sce_integrated)$Symbol]
# hsps <- hsps[hsps %in% rowData(sce_fvc)$Symbol]
```

# Single run of a Pseudobulk DE


```{r manual 1}
# Fresh Sole_Boldo_et_al He_et_al Tabib_et_al
control_protocol <- "Fresh"
test_protocol <- "Culture"

refactorise <- function(sce_object, column_vector){
  if (length(column_vector) == 1){
    colData(sce_object)[,column_vector] <- factor(colData(sce_object)[, column_vector])
  } else {
    for (col in column_vector){
      colData(sce_object)[,col] <- factor(colData(sce_object)[,col])
    }
  }
  sce_object
}

sce_subset <- sce_fvc[,sce_fvc$Protocol %in% c(control_protocol, test_protocol)]
sce_subset <- refactorise(sce_subset, c("Protocol", "Sample", "manual_labels_coarse"))

sce_muscat <- prepSCE(sce_subset,
                      kid="manual_labels_coarse",
                      gid="Protocol",
                      sid="Sample",
                      drop=TRUE)

# table(sce_subset$Sample, sce_subset$manual_labels_coarse)
```

We store number of clusters and number of sample IDs as variables

```{r manual 2}
nk <- length(kids <- levels(sce_muscat$cluster_id))
ns <- length(sids <- levels(sce_muscat$sample_id))
names(kids) <- kids; names(sids) <- sids

pb <- aggregateData(sce_muscat, assay="counts", fun="sum", by=c("cluster_id", "sample_id"))

(pb_mds <- pbMDS(pb))
```

```{r, include=FALSE}
# res1 <- pbDS(pb)
# sapply(levels(res$sce$cluster_id), function(cell_type){
#   print(cell_type)
#   pbHeatmap(res$sce, res$muscat, k = cell_type)
# })

# pbHeatmap(sce_muscat, res1, top_n = 2, fdr = 1, lfc = 0)

# library(limma)
# ei <- metadata(sce_muscat)$experiment_info
# mm <- model.matrix(~ ei$group_id)
# dimnames(mm) <- list(ei$sample_id, levels(ei$group_id))
# contrast <- makeContrasts("Sole_Boldo_et_al-Fresh", levels = mm)

# res1 <- pbDS(pb)
# res2 <- pbDS(pb, design = mm, contrast = contrast)
# res1==res2
# 
# # Are they the same?
# res1$table[[1]][[1]]

# tbl <- res$table[[1]]
# names(tbl)
# k1 <- tbl[[1]]
```


# Run DS for all Protocols

```{r ds functions}
run_ds <- function(sce_object, control_protocol, test_protocol){
  sce_subset <- sce_object[,sce_object$Protocol %in% c(control_protocol, test_protocol)]
  sce_subset <- refactorise(sce_subset, c("Protocol", "Sample", "manual_labels_coarse"))
  
  sce_muscat <- prepSCE(sce_subset,
                        kid="manual_labels_coarse",
                        gid="Protocol",
                        sid="Sample",
                        drop=TRUE)
  
  nk <- length(kids <- levels(sce_muscat$cluster_id))
  ns <- length(sids <- levels(sce_muscat$sample_id))
  names(kids) <- kids; names(sids) <- sids
  
  pb <- aggregateData(sce_muscat, assay="counts", fun="sum", by=c("cluster_id", "sample_id"), verbose=F)
  res <- pbDS(pb, verbose=F)
  list("muscat"=res, "sce"=sce_muscat)
}

print_tables <- function(res){
  sapply(names(res$muscat$table[[1]]), function(cell_type){
    cat(paste0("\n\n#### ", cell_type," \n\n"))
    print(kable_styling(kable(head(format(res$muscat$table[[1]][[cell_type]], digits=2), 10))))
  })
}

print_hsp_tables <- function(res){
  sapply(names(res$muscat$table[[1]]), function(cell_type){
    tbl <- res$muscat$table[[1]][[cell_type]]
    hsps_in_table <- hsps[hsps %in% tbl$gene]
    cat(paste0("\n\n#### ", cell_type," \n\n"))
    tbl <- tbl[tbl$gene %in% hsps_in_table,] %>%
      arrange(p_adj.loc)
    print(kable_styling(kable(format(tbl, digits=2))))
  })
}

plot_topn_heatmap <- function(res, title){
  as.ggplot(pbHeatmap(res$sce, res$muscat, top_n=20)) +
    labs(title=title)
}

plot_hsp_heatmap <- function(res, title){
  hsps_in_dataset <- hsps[hsps %in% rownames(res$sce)]
  as.ggplot(pbHeatmap(res$sce, res$muscat, g = hsps_in_dataset, fdr=1, lfc=0)) +
    labs(title=title)
}

# Run DE analyses
res_tabib <- run_ds(sce_integrated, "Fresh", "Tabib_et_al")
res_sb <- run_ds(sce_integrated, "Fresh", "Sole_Boldo_et_al")
res_he <- run_ds(sce_integrated, "Fresh", "He_et_al")
res_tabib_v_sb <- run_ds(sce_integrated, "Tabib_et_al", "Sole_Boldo_et_al")
res_tabib_v_he <- run_ds(sce_integrated, "Tabib_et_al", "He_et_al")
res_sb_v_he <- run_ds(sce_integrated, "Sole_Boldo_et_al", "He_et_al")
res_culture <- run_ds(sce_fvc, "Fresh", "Culture")
```

# Top DS genes

```{r table top, results="asis"}
cat("\n\n## Top DS genes: tabular data {.tabset} \n\n")
cat("\n\n### Fresh vs. Tabib et al {.tabset} \n\n")
invisible(print_tables(res_tabib))
cat("\n\n### Fresh vs. He et al {.tabset}\n\n")
invisible(print_tables(res_he))
cat("\n\n### Fresh vs. Sole-Boldo et al  {.tabset} \n\n")
invisible(print_tables(res_sb))
cat("\n\n### Fresh vs. Cultured  {.tabset} \n\n")
invisible(print_tables(res_culture))

cat("\n\n### Tabib vs. Sole-Boldo  {.tabset} \n\n")
invisible(print_tables(res_tabib_v_sb))
cat("\n\n### Tabib vs. He  {.tabset} \n\n")
invisible(print_tables(res_tabib_v_he))
cat("\n\n### Sole-Boldo vs. He  {.tabset} \n\n")
invisible(print_tables(res_sb_v_he))
```


```{r heatmap top, results="asis", fig.width=7, fig.height=12}
cat("\n\n## Top DS genes: heatmap {.tabset} \n\n")
cat("\n\n### Fresh vs. Tabib et al \n\n")
plot_topn_heatmap(res_tabib, "Tabib et al")
cat("\n\n### Fresh vs. He et al \n\n")
plot_topn_heatmap(res_he, "He et al")
cat("\n\n### Fresh vs. Sole-Boldo et al \n\n")
plot_topn_heatmap(res_sb, "Sole-Boldo et al")
cat("\n\n### Fresh vs. Cultured  {.tabset} \n\n")
plot_topn_heatmap(res_culture, "Cultured")

cat("\n\n### Tabib vs. Sole-Boldo  {.tabset} \n\n")
plot_topn_heatmap(res_tabib_v_sb, "Tabib vs. Sole-Boldo")
cat("\n\n### Tabib vs. He  {.tabset} \n\n")
plot_topn_heatmap(res_tabib_v_he, "Tabib vs. He")
cat("\n\n### Sole-Boldo vs. He  {.tabset} \n\n")
plot_topn_heatmap(res_sb_v_he, "Sole-Boldo vs. He")
```

# HSP DS

```{r table hsp, results="asis"}
cat("\n\n## HSP genes: tabular data {.tabset} \n\n")
cat("\n\n### Fresh vs. Tabib et al {.tabset} \n\n")
invisible(print_hsp_tables(res_tabib))
cat("\n\n### Fresh vs. He et al {.tabset} \n\n")
invisible(print_hsp_tables(res_he))
cat("\n\n### Fresh vs. Sole-Boldo et al {.tabset} \n\n")
invisible(print_hsp_tables(res_sb))
cat("\n\n### Fresh vs. Cultured  {.tabset} \n\n")
invisible(print_hsp_tables(res_culture))

cat("\n\n### Tabib vs. Sole-Boldo  {.tabset} \n\n")
invisible(print_hsp_tables(res_tabib_v_sb))
cat("\n\n### Tabib vs. He  {.tabset} \n\n")
invisible(print_hsp_tables(res_tabib_v_he))
cat("\n\n### Sole-Boldo vs. He  {.tabset} \n\n")
invisible(print_hsp_tables(res_sb_v_he))
```

```{r heatmap hsp, results="asis", fig.width=7, fig.height=12}
cat("\n\n## HSP genes: heatmaps {.tabset} \n\n")
cat("\n\n### Fresh vs. Tabib et al \n\n")
plot_hsp_heatmap(res_tabib, "Fresh vs. Tabib et al")
cat("\n\n### Fresh vs. He et al \n\n")
plot_hsp_heatmap(res_he, "Fresh vs. He et al")
cat("\n\n### Fresh vs. Sole-Boldo et al \n\n")
plot_hsp_heatmap(res_sb, "Fresh vs. Sole-Boldo et al")
cat("\n\n### Fresh vs. Cultured  {.tabset} \n\n")
plot_hsp_heatmap(res_culture, "Fresh vs. Cultured")

cat("\n\n### Tabib vs. Sole-Boldo  {.tabset} \n\n")
plot_hsp_heatmap(res_tabib_v_sb, "Tabib vs. Sole-Boldo")
cat("\n\n### Tabib vs. He  {.tabset} \n\n")
plot_hsp_heatmap(res_tabib_v_he, "Tabib vs. He")
cat("\n\n### Sole-Boldo vs. He  {.tabset} \n\n")
plot_hsp_heatmap(res_sb_v_he, "Sole-Boldo vs. He")
```



```{r}
sessionInfo()
```
