---
title: "Fresh vs. Culture: Annotations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressMessages({
  library(hrbrthemes)
  library(scater)
  library(scran)
  library(ggplot2)
  library(SingleCellExperiment)
  library(plyr)
  library(dplyr)
  library(RColorBrewer)
  library(openxlsx)
  library(ggplotify)
  library(pheatmap)
  library(kableExtra)
  library(viridis)
  library(parallel)
})

n_workers <- min(detectCores()-1, 16)
RhpcBLASctl::blas_set_num_threads(n_workers)
bpparam <- BiocParallel::MulticoreParam(workers=n_workers, RNGseed = 123)

source("../code/general_purpose_code.R")
```

# Load data

We load the data for the two 'Fresh' samples and four SSc skin samples that we have processed in the previous script.  All six samples were jointly batch-corrected.

```{r load-data}
sce <- readRDS(file="../data/EOS_files/FvC_EOS2.rds")
```


# Load and add labels for clusters

```{r load csv with cluster annotations}
cluster_annotation_mapping <- read.xlsx("../metadata/cluster_annotation_map.xlsx", sheet="Fresh_v_control")
kableExtra::kable(cluster_annotation_mapping)

sce$manual_labels_coarse <- mapvalues(sce$label, cluster_annotation_mapping$Cluster.ID, cluster_annotation_mapping$Annotation)
```

# UMAPs coloured by cluster and cell type group {.tabset}

We take a look at the UMAPs to check that the mapping is indeed as we expect.

```{r visualise UMAPs, fig.width=16, fig.height=10, results="asis"}
cat("\n\n## Unsupervised clustering \n\n")
plotReducedDim(sce, "UMAP", colour_by="label", text_by="label", point_size=0.25) +
  labs(title="Unsupervised clustering",
       subtitle="Fresh and cultured samples") +
  theme_ipsum_rc() +
  theme(legend.position="bottom",
        panel.spacing = unit(5, "mm")) +
  guides(color=guide_legend(title="",
                            nrow=1,
                            byrow=TRUE, 
                            override.aes = list(size=4, shape = 15, alpha = 1)))

cat("\n\n## Cell type groups \n\n")
fig9a1 <- plotReducedDim(sce, "UMAP", colour_by="manual_labels_coarse", text_by="manual_labels_coarse", point_size=0.25) +
  labs(title="Cell annotation",
       subtitle="Fresh and cultured samples") +
  theme_ipsum_rc() +
  get_cell_colours(levels(sce$manual_labels_coarse)) +
  theme(#legend.position="bottom",
        panel.spacing = unit(5, "mm")) +
  guides(color=guide_legend(title="",
                            #nrow=1,
                            byrow=TRUE, 
                            override.aes = list(size=4, shape = 15, alpha = 1)))
fig9a1

fig9a2 <- plotReducedDim(sce, "UMAP", colour_by="Protocol", point_size=0.25) +
  labs(title="Protocol",
       subtitle="Fresh and cultured samples") +
  theme_ipsum_rc() +
  get_protocol_colours(levels(sce$Protocol)) +
  theme(#legend.position="bottom",
        panel.spacing = unit(5, "mm")) +
  guides(color=guide_legend(title="",
                            #nrow=1,
                            byrow=TRUE, 
                            override.aes = list(size=4, shape = 15, alpha = 1)))
fig9a2
```




# Frequency of cell types

```{r, fig.width=14, fig.height=10}
ret_vals <- plot_label_barplots(sce, "manual_labels_coarse", level1="Sample", level2=NA, print=F, add_labels=c(FALSE, TRUE))

fig9b1 <- ret_vals$plt_rel +
  get_cell_colours(sce$manual_labels_coarse, type="fill") +
  theme_ipsum_rc() +
  theme(strip.text.y = element_blank(),
        strip.text.x = element_text(size = 10),
        axis.text.x=element_blank(),
        panel.spacing = unit(5, "mm"),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.5,  # Bottom margin
                             l = 20)) +# Left margin
  guides(fill=guide_legend(title="Cell type")) +
  labs(title="Cell type frequency",
       subtitle="Fresh and cultured samples",
       y="",
       x="")
fig9b1

fig9b2 <- ret_vals$plt_abs +
  get_cell_colours(sce$manual_labels_coarse) +
  scale_y_continuous(n.breaks=3) +
  get_cell_colours(sce$manual_labels_coarse, type="fill") +
  theme_ipsum_rc() +
  theme(strip.text.y = element_blank(),
        strip.text.x = element_text(size = 10),
        axis.text.x=element_blank(),
        panel.spacing = unit(5, "mm"),
        plot.margin = margin(t = 0.5,  # Top margin
                             r = 0.5,  # Right margin
                             b = 0.5,  # Bottom margin
                             l = 20)) # Left margin 
fig9b2
```

```{r plot abundances v2, fig.width=20, fig.height=10}
count_table <- table(sce$manual_labels_coarse, sce$Sample)
lib_sizes <- colSums(count_table)

sample_protocol_mapping <- as.data.frame(colData(sce)) %>%
  group_by(Protocol, Sample) %>%
  summarise() %>%
  dplyr::ungroup() %>%
  tibble::column_to_rownames('Sample')



fig9c <- purrr::map(unique(sce$manual_labels_coarse), ~ {
  tibble::tibble(counts = count_table[.x,],
                 abundances = counts/lib_sizes,
                 Protocol= sample_protocol_mapping[levels(sce$Sample), "Protocol"],
                 Cluster = .x)}) %>% 
  purrr::reduce(rbind) %>% 
  dplyr::arrange("Cluster") %>%
  ggplot() + 
  geom_boxplot(aes(y=abundances,x=Protocol, color=Protocol)) +
  geom_point(aes(y=abundances,x=Protocol, color=Protocol), size = 3) +
  theme_ipsum_rc() +
  theme(axis.text.x = element_text(angle=45, hjust = 1), strip.text.x = element_text(family = "Helvetica",face = "bold"),
        legend.position="none") +
  scale_y_continuous(labels = scales::percent) +
  labs(title="Relative cell type abundance", x="", y="") +
  facet_wrap(~Cluster, scales = "free_y",ncol = 6, nrow = 2) +
  get_protocol_colours(levels(sce$Protocol))
fig9c
```

# Figure 9.1

```{r figure 9.1, fig.width=10, fig.height=12}
folder_path <- "/Users/dominiquepaul/xRobinson Group/01_Systemic Sclerosis/output/plots/"

figure9_1 <- ggpubr::ggarrange(
  fig9a1,
  fig9a2,
  ncol = 1, nrow = 2,labels = "AUTO"
)

suppressWarnings(figure9_1)

ggsave(paste0(folder_path, "figure9_1.png"),
       plot=figure9_1,
       width=20*0.75,
       height=24*0.75,
       units="in",
       bg="white")
```

# Figure 9.2

```{r figure 9.2, fig.width=20, fig.height=24}
figure9_2 <- ggpubr::ggarrange(
  ggpubr::ggarrange(fig9b1, fig9b2, ncol = 2),
  fig9c,
  ncol = 1, nrow = 2,labels = "AUTO",
  heights=c(1.5,1)
)

suppressWarnings(figure9_2)

ggsave(paste0(folder_path, "figure9_2.png"),
       plot=figure9_2,
       width=20*0.9,
       height=24*0.9,
       units="in",
       bg="white")
```



```{r Fresh_v_culture supplemental, fig.width=16, fig.height=10}
FVC_supplemental <- plotReducedDim(sce, "UMAP", colour_by="label", text_by="label", point_size=0.25, text_size=4) +
  labs(title="Unsupervised clustering",
       subtitle = "Fresh and cultured samples") +
  theme_ipsum_rc() +
  # get_cell_colours(levels(sce$manual_labels_coarse)) +
  # theme(legend.position="bottom") +
  guides(color=guide_legend(title="",
                            # nrow=3,
                            byrow=TRUE, 
                            override.aes = list(size=4, shape = 15, alpha = 1)))

FVC_supplemental

ggsave(paste0(folder_path, "FVC_supplemental.png"),
       plot=FVC_supplemental,
       width=16*0.9,
       height=10*0.9,
       units="in",
       bg="white")

```


<!-- # Cell type to sample -->

<!-- ```{r cell type to sample table, fig.height=6, fig.width=10} -->
<!-- tab <- table(sce$manual_labels_coarse, sce$Sample) -->
<!--   tab_annotation <- tab -->
<!--   tab_annotation[tab_annotation==0] <- "" -->

<!--   as.ggplot(pheatmap(tab, color=brewer.pal(n=9, name="Oranges"), display_numbers=tab_annotation, silent=T)) + -->
<!--     theme_ipsum_rc() + -->
<!--     labs(title="Clusters and samples") + -->
<!--     theme(axis.line=element_blank(),axis.text.x=element_blank(), -->
<!--             axis.text.y=element_blank(),axis.ticks=element_blank(), -->
<!--             axis.title.x=element_blank(), -->
<!--             axis.title.y=element_blank(),legend.position="none", -->
<!--             panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(), -->
<!--             panel.grid.minor=element_blank(),plot.background=element_blank()) -->
<!-- ``` -->



<!-- # Marker genes {.tabset} -->

<!-- Marker genes can help us identify clusters as specific cell types and also discover smaller changes between subclusters (e.g., changes in activation or differentiation state) -->

<!-- ```{r prepare sheet to save data to} -->
<!-- wb <- createWorkbook() -->
<!-- addWorksheet(wb, "Intro") -->
<!-- writeData(wb, "Intro", data.frame("Marker genes for the protocol comparison"), colNames=FALSE) -->
<!-- ``` -->

<!-- ```{r function for top_genes} -->
<!-- plot_heatmap_of_top_genes <- function(sce, markers, top_n=5){ -->
<!--   # function assumes the sce columns: label, Sample -->

<!--   # get top genes -->
<!--   top_genes <- lapply(markers, function(marker_set){ -->
<!--     marker_set %>% -->
<!--       as.data.frame() %>% -->
<!--       filter(Top <= top_n) %>% -->
<!--       rownames() -->
<!--   }) %>% -->
<!--     unlist(use.names=F) %>% -->
<!--     unique() -->

<!--   # Matrix to plot the heatmap of expression values. -->
<!--   # To avoid a too big plot we only keep a fraction of the cells i there are too many -->
<!--   nth_cell <- max(1, round(dim(sce)[2] / 1000,0)) -->
<!--   sce_plot <- sce[top_genes, seq(1, dim(sce)[2], nth_cell)] -->
<!--   # We sort by cluster -->
<!--   sce_plot <- sce_plot[,order(sce_plot$label, sce_plot$Sample)] -->
<!--   # Anotations -->
<!--   column_annotations <- data.frame("Sample"=sce_plot$Sample, "Cluster"=factor(sce_plot$label), row.names=colnames(sce_plot)) -->
<!--   # clip heatmap values to ensure that colour contrast is high -->
<!--   sce_plot <- assay(sce_plot, "logcounts") -->
<!--   sce_plot[sce_plot >=5] = 5 -->

<!--   print(as.ggplot(pheatmap(sce_plot, annotation_col=column_annotations, -->
<!--            show_rownames=TRUE, show_colnames=FALSE, cluster_cols=FALSE, silent=TRUE, -->
<!--            color=inferno(50))) + -->
<!--     labs(title="Gene signatures per cluster identified", -->
<!--          subtitle=paste0("Top ", top_n, " differentially expressed genes per cluster; values are log-counts clipped at 5")) + -->
<!--     theme_ipsum_rc()+ -->
<!--     theme(axis.line=element_blank(),axis.text.x=element_blank(), -->
<!--             axis.text.y=element_blank(),axis.ticks=element_blank(), -->
<!--             axis.title.x=element_blank(), -->
<!--             axis.title.y=element_blank(),legend.position="none", -->
<!--             panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(), -->
<!--             panel.grid.minor=element_blank(),plot.background=element_blank())) -->
<!-- } -->
<!-- ``` -->

<!-- ```{r marker genes, fig.height=12, fig.width=6, results="asis"} -->

<!-- sce_markers <- sce -->
<!-- colLabels(sce_markers) <- factor(sce_markers$manual_labels_coarse) -->

<!-- # we filter for only the highly variable genes. Otherwise we will see many genes which -->
<!-- # are *significantly* different (high p-values), but not marginally different (low logFCs) -->
<!-- markers.all.up <- findMarkers(sce_markers[rowData(sce_markers)$is_hvg,], -->
<!--                               direction="up", -->
<!--                               lfc=1, -->
<!--                               block=sce_markers$Sample) -->

<!-- for (i in 1:length(unique(names(markers.all.up)))){ -->

<!--   cell_type_of_cluster <- unique(names(markers.all.up))[[i]] -->
<!--   cat("\n\n### ", cell_type_of_cluster,  "cluster \n\n") -->

<!--   chosen <- i -->

<!--   interesting.all.up <- markers.all.up[[chosen]] -->

<!--   print(kable_styling(kable(interesting.all.up[1:10,], digits=2))) -->
<!--   cat("\n") -->

<!--   best.set <- interesting.all.up[1:50,] -->
<!--   logFCs <- getMarkerEffects(best.set) -->

<!--   # write to excel -->
<!--   sheet_name <- paste0("All_data", "_ttest_pairs_cl_", i) -->
<!--   addWorksheet(wb, sheet_name) -->
<!--   writeData(wb, sheet_name, logFCs, rowNames=TRUE) -->

<!--   # set all NAs to 0 -->
<!--   cat(paste0(sum(is.na(logFCs)), " NA values in cluster ", chosen, "\n")) -->
<!--   logFCs[is.na(logFCs)] <- 0 -->

<!--   pheatmap(logFCs, breaks=seq(-5, 5, length.out=101), main=paste0("Cluster ",i, ": Marker genes, corrected for batch effects \n (upregulated only, pval.type='all')")) -->
<!-- } -->
<!-- ``` -->

<!-- Averaged expression values for markers -->

<!-- ```{r averaged expression values for markers, fig.width=14, fig.height=35, results="asis"} -->
<!-- plot_heatmap_of_top_genes(sce_markers, markers.all.up, top_n=5) -->
<!-- ``` -->


<!-- # Marker genes identified by entropy & f-stat -->

<!-- The following shows the marker genes as identified via the entropy-fstat-method for the dataset combining all samples as well as the dataset for all samples combined. -->

<!-- ```{r entropy f-stat, results="asis", fig.width=10, fig.height=7, results="asis", warning=FALSE, eval=F} -->

<!-- sce_hvg_subset <- sce[rowData(sce)$is_hvg,] -->

<!-- # sce_object <- sce_object[rowData(sce_object)$is_hvg==TRUE,] -->

<!-- shannon_entropy <- function(x, e = 1e-10) { -->
<!--   x <- x+e # allow zeros in x -->
<!--   p <- (x)/sum(x) -->
<!--   -sum((p)*log(p)) -->
<!-- } -->

<!-- aggr_syn_sce <- aggregateAcrossCells(sce_hvg_subset, sce_hvg_subset$label) -->

<!-- entropy_fstat_sce <- BiocParallel::bplapply( -->
<!--   seq_along(rownames(aggr_syn_sce)), -->
<!--   function(.x) { -->
<!--     data.frame( -->
<!--       gene = rowData(sce_hvg_subset)$Symbol[.x], -->
<!--       entropy = shannon_entropy(counts(aggr_syn_sce[.x,])), -->
<!--       fstat = summary(aov(logcounts(sce_hvg_subset)[.x, ] ~ colLabels(sce_hvg_subset)))[[1]]$`F value`[1] -->
<!--     ) -->
<!--     }, BPPARAM = BiocParallel::MulticoreParam(workers=n_workers, RNGseed = 123)) %>% -->
<!--   purrr::reduce(rbind) -->

<!-- entropy_fstat_sce <- entropy_fstat_sce %>% -->
<!--   dplyr::mutate(entropy_norm = (entropy+1e-3)/(max(entropy)+1e-3), -->
<!--                 fstat_norm = fstat/max(fstat,na.rm = TRUE), -->
<!--                 ent_fstat_ratio = fstat_norm / entropy_norm, na.rm = TRUE) -->

<!-- topmarkers <- order(entropy_fstat_sce$ent_fstat_ratio, decreasing = TRUE)[1:50] -->

<!-- topmarker_knownmarkers <- unique(entropy_fstat_sce[topmarkers,]) -->

<!-- cat("\n\n### Entropy-fstat plot \n\n") -->
<!-- print(ggplot(entropy_fstat_sce) + -->
<!--   geom_point(aes(x=log(fstat), y=entropy), show.legend = F, colour="#02C39A") + -->
<!--   labs(x="log(F-statistic)", y="Entropy", main="Entropy vs. F-statistic of cluster separation") + -->
<!--   ggrepel::geom_text_repel(data=topmarker_knownmarkers, -->
<!--                            aes(x=log(fstat), y=entropy, label=gene, size=4), -->
<!--                            show.legend=F, -->
<!--                            max.overlaps = 30)) -->

<!-- # write to excel -->
<!-- sheet_name <- paste0("All_data", "_fstat_top_markers") -->
<!-- addWorksheet(wb, sheet_name) -->
<!-- writeData(wb, sheet_name, topmarker_knownmarkers, rowNames=TRUE) -->

<!-- # Expression plots -->
<!-- cat("\n\n### Expression plots for all samples combined {.tabset} \n\n") -->

<!-- genes_to_plot <-  unique(entropy_fstat_sce$gene[topmarkers]) -->
<!-- rownames(sce_hvg_subset) <- rowData(sce_hvg_subset)$Symbol -->

<!-- for(i in seq_along(genes_to_plot)){ -->
<!--   cat("\n\n#### ", genes_to_plot[i], "\n\n") -->
<!--   print(plotExpression(sce_hvg_subset,   -->
<!--                  entropy_fstat_sce$gene[entropy_fstat_sce$gene == genes_to_plot[i]], -->
<!--                  x = "label", colour_by = "Sample", point_size=0.75)) -->
<!-- } -->
<!-- ``` -->


<!-- ```{r save marker genes to excel, eval=F} -->
<!-- # save results -->
<!-- filename_marker_results <- "../docs/output/Protocol_analysis_custom_labels_markers_results_hvg_genes.xlsx" -->
<!-- saveWorkbook(wb, file = filename_marker_results, overwrite = TRUE) -->
<!-- ``` -->


# Save SCE object

```{r save sce object, class.source="fold-show"}
file_path <- "../data/EOS_files/FvC_EOS3.rds"
saveRDS(sce, file=file_path)
upload_file_to_iSEE(file_path)
```

