---
title: "Fresh vs. Culture: Gene Expression Plot and DA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressMessages({
  library(hrbrthemes)
  library(ggplot2)
  library(SingleCellExperiment)
  library(plyr)
  library(dplyr)
  library(openxlsx)
  library(scater)
  library(edgeR)
  library(RColorBrewer)
  library(viridis)
})

source("../code/general_purpose_code.R")
``` 

# Load data

We load the data for the two 'Fresh' samples and four SSc skin samples that we have processed in the previous script.  All six samples were jointly batch-corrected.

```{r load-data}
sce <- readRDS(file="../data/EOS_files/FvC_EOS3_sce.rds")

sce$manual_labels_coarse <- sub("Vascular endothelial cells", "Vascular endothelial", sce$manual_labels_coarse)
sce$manual_labels_coarse <- factor(sce$manual_labels_coarse)
unique(sce$manual_labels_coarse)
```


# Plot comparing expression values for selected genes

Frac_cells_expressed_in = Number of e.g. Keratinocyte cells in sample X with an expression value > 0 for the given gene / Number of keratinocyte cells in sample X

Expr_avg_to_overall_avg = Average expression value of gene in e.g. keratinocytes of Sample X  


```{r read excel with markers}
marker_overview <- read.xlsx(xlsxFile="../Metadata/Special_plot_markers.xlsx")
```


```{r create plot, fig.width=20, fig.height=7}
# create initial data frame
genes_ball_plot <- marker_overview %>%
  dplyr::rename(cell_type=Cell.type, gene=Marker)

cat("Genes not found in data: ")
cat( genes_ball_plot$gene[!genes_ball_plot$gene %in% rownames(sce)])
genes_ball_plot <- genes_ball_plot[genes_ball_plot$gene %in% rownames(sce),]


genes_ball_plot_with_values <- lapply(unique(sce$Sample), function(sample){
  df_intermediate_level1 <- genes_ball_plot
  # add the sample name as a column
  df_intermediate_level1$sample <- sample
  
  # for loop that uses only the subset of the sce for the cell type
  df_intermediate_level1 <- lapply(unique(genes_ball_plot$cell_type), function(cell_type){
    cat(paste0("    ", cell_type, "\n\n"))
    genes_for_cell_type <- genes_ball_plot[genes_ball_plot$cell_type == cell_type, "gene"]
    df_intermediate_level2 <- df_intermediate_level1[df_intermediate_level1$cell_type == cell_type,]
    sce_subset <- sce[,sce$Sample == sample & sce$manual_labels_coarse == cell_type]
    sce_anti_subset <- sce[,sce$Sample == sample & sce$manual_labels_coarse != cell_type]
    # check whether the genes really occur in the sample
    mask <- genes_for_cell_type %in% rownames(sce_subset)
    df_intermediate_level2$mean_of_logs_cell_type_only <- 0
    df_intermediate_level2$frac_cells_expressed_in_cell_type_only <- 0
    
    if (dim(sce_subset)[2] > 0){ # check that the sample is not empty
      if (length(genes_for_cell_type) > 1){ # if there is only one gene, then the rowMeans function needs to be replaced by mean
        if (dim(sce_subset)[2] > 1){ # if there is only one cell then we dont take averages
          df_intermediate_level2[mask, "mean_of_logs_cell_type_only"] <- rowMeans(logcounts(sce_subset)[genes_for_cell_type[mask],])
          df_intermediate_level2[mask, "frac_cells_expressed_in_cell_type_only"] <- rowMeans(assay(sce_subset)[genes_for_cell_type[mask],] > 0)
        } else {
          cat(paste0("Warning: ", sample, " only has one cell of type ", cell_type))
          df_intermediate_level2[mask, "mean_of_logs_cell_type_only"] <- logcounts(sce_subset)[genes_for_cell_type[mask],]
          df_intermediate_level2[mask, "frac_cells_expressed_in_cell_type_only"] <- as.numeric(assay(sce_subset)[genes_for_cell_type[mask],] > 0)
        }
      } else {
        df_intermediate_level2[mask, "mean_of_logs_cell_type_only"] <- mean(logcounts(sce_subset)[genes_for_cell_type[mask],])
        df_intermediate_level2[mask, "frac_cells_expressed_in_cell_type_only"] <- mean(assay(sce_subset)[genes_for_cell_type[mask],] > 0)
      }
    }
    
    if (length(genes_for_cell_type) > 1){ # if there is only one gene, then the rowMeans function needs to be replaced by mean
      df_intermediate_level2$avg_expression_other_cells <- rowMeans(logcounts(sce_anti_subset)[genes_for_cell_type,])
    } else {
      df_intermediate_level2$avg_expression_other_cells <- mean(logcounts(sce_anti_subset)[genes_for_cell_type,])
    }
    
    df_intermediate_level2
  }) %>% ldply(data.frame)

  df_intermediate_level1
  
}) %>% ldply(data.frame)


# add relative column
genes_ball_plot_with_values$expr_avg_to_overall_avg <- genes_ball_plot_with_values$mean_of_logs_cell_type_only / genes_ball_plot_with_values$avg_expression_other_cells
genes_ball_plot_with_values[is.na(genes_ball_plot_with_values$expr_avg_to_overall_avg), "expr_avg_to_overall_avg"] <- 0
genes_ball_plot_with_values[genes_ball_plot_with_values$expr_avg_to_overall_avg > 50, "expr_avg_to_overall_avg"] <- 50

# add protocol as a column
sample2protocol <- data.frame("sample"=levels(colData(sce)$Sample)) %>%
  mutate("Protocol"=sub("_S[0-9]*", "", sample)) %>%
  mutate("Protocol"=sub("Skin","SSc", Protocol))
  
genes_ball_plot_with_values <- genes_ball_plot_with_values %>%
  left_join(sample2protocol)

# use this to normalise the colour scale
# col_norm_const <- max(genes_ball_plot_with_values$mean_of_logs_cell_type_only )

# fix cell type column as factor to ensure correct ordering
genes_ball_plot_with_values$cell_type <- factor(genes_ball_plot_with_values$cell_type, levels=unique(genes_ball_plot$cell_type))

```

```{r plot and save, fig.width=20, fig.height=7}
folder_path <- "/Users/dominiquepaul/xRobinson Group/01_Systemic Sclerosis/output/plots/"

# Change order of Sample levels
genes_ball_plot_with_values$sample <- factor(genes_ball_plot_with_values$sample, levels=c("Culture_S5", "Culture_S4", "Culture_S2", "Culture_S1","Fresh_S2", "Fresh_S1"))
genes_ball_plot_with_values$Protocol <- factor(genes_ball_plot_with_values$Protocol, levels=c("Fresh", "Culture"))

# rename the levels to have shorter plot panel titles
levels(genes_ball_plot_with_values$cell_type) <- sub("Vascular endothelial", "Vasc. Endo.", levels(genes_ball_plot_with_values$cell_type))
levels(genes_ball_plot_with_values$cell_type) <- sub("Lymphatic endothelial", "Lym. Endo.", levels(genes_ball_plot_with_values$cell_type))
levels(genes_ball_plot_with_values$cell_type) <- sub("Sweat gland cells", "Sweat gland", levels(genes_ball_plot_with_values$cell_type))
levels(genes_ball_plot_with_values$cell_type) <- sub("Dendritic cells", "Dendritic", levels(genes_ball_plot_with_values$cell_type))
levels(genes_ball_plot_with_values$cell_type) <- sub("Schwann cells", "Schwann", levels(genes_ball_plot_with_values$cell_type))
#levels(genes_ball_plot_with_values$cell_type) <- sub("Melanocytes", "Sweat gland.", levels(genes_ball_plot_with_values$cell_type))


special_plot <- ggplot(genes_ball_plot_with_values, aes(x=gene, y=sample, size=frac_cells_expressed_in_cell_type_only, color=mean_of_logs_cell_type_only), otherInfo(Protocol)) +
  geom_point() +
  facet_grid(Protocol~cell_type, scales="free", space="free") +
  theme_ipsum_rc() +
  theme(axis.text.x = element_text(angle = 90, hjust=1), 
        axis.ticks.x=element_blank(),
        panel.spacing = unit(3, "mm"),
        strip.text.y = element_blank(),
        strip.text.x = element_text(size=11)
        ) +
  labs(title="Comparison of selected genes",
       subtitle="Size: proportion of cells of each sample expressing the marker \nColour:mean of log expression",
       y="", x="") +
  scale_colour_viridis() +
  guides(color=guide_colorbar(title=c("Mean of log counts")),
         size=guide_legend(title=c("% of cell expressed in cell type"))) 
  
special_plot

ggsave(paste0(folder_path, "special_plot.png"),
       plot=special_plot,
       width=20*0.9,
       height=7*0.9,
       units="in",
       bg="white")
```
