---
title: "Cell Cycle Analysis"
author: "Dominique Paul"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

suppressMessages({
  library(RCurl)
  library(Seurat)
  library(hrbrthemes)
  library(ggplot2)
  library(SingleCellExperiment)
  library(dplyr)
  library(scater)
})

theme_set(theme_ipsum_rc())

# this script contains the function 'plot_stratified_views' which makes plotting 
# dimension reductions easier and is used throughout the project
source("../code/general_purpose_code.R")
```

# Load data

We load the data for the two 'Fresh' samples as well as the joint dataset of fresh AND cultured samples that we have processed in the previous script.  Both the two and the six samples were batch-corrected respectively.

```{r load data}
file_path_fresh <- "../data/EOS_files/Fresh_only_EOS1.rds"
sce_fresh <- readRDS(file_path_fresh)
rownames(sce_fresh) <- rowData(sce_fresh)$ID # might need this for the analysis

file_path_all <- "../data/EOS_files/FvC_EOS1.rds"
sce_all <- readRDS(file_path_all)
rownames(sce_all) <- rowData(sce_all)$ID # might need this for the analysis
```

# Download cell cycle genes

We download the marker genes for the cell cycle stages from a github repository. The data comes in a data frame. We split the genes into lists according to the cell cycle phase.

```{r get cell cycle genes}
cc_file <- getURL("https://raw.githubusercontent.com/hbc/tinyatlas/master/cell_cycle/Homo_sapiens.csv") 
cell_cycle_genes <- read.csv(text = cc_file)

head(cell_cycle_genes)

s_genes <- cell_cycle_genes %>%
  dplyr::filter(phase == "S") %>%
  pull("geneID")

g2m_genes <- cell_cycle_genes %>%
  dplyr::filter(phase == "G2/M") %>%
  pull("geneID")
```


# Identify cell cycle phases

The function we are using for the identification of the cell cycle is a Seurat function. Seurat is a data format similar to the Single Cell Experiment object we generally use in the other scrips of this analysis. So we convert our Single Cell Experiment object into a Seurat object and run the identification of the cell cycle phase. We save the results to our Single Cell Experiment object as we will be using this object class it in our remaining analysis.


```{r add cell cycle phase, warning=FALSE}
## Fresh samples only
sce_seurat_fresh <- as.Seurat(sce_fresh, counts="counts", data="logcounts")

sce_seurat_fresh <- CellCycleScoring(sce_seurat_fresh,
                               g2m.features = g2m_genes,
                               s.features = s_genes)

sce_fresh$phase <- sce_seurat_fresh$Phase

## Fresh and cultured samples
sce_seurat_all <- as.Seurat(sce_all, counts="counts", data="logcounts")

sce_seurat_all <- CellCycleScoring(sce_seurat_all,
                               g2m.features = g2m_genes,
                               s.features = s_genes)

sce_all$phase <- sce_seurat_all$Phase
```

# Visualisation

## Coloured by clusters label {.tabset}

We show the plots once to inspect the clusters. It could be interesting to see that certain clusters might have been formed based on the similarity of the cell cycle phase.

```{r, fig.width=12, fig.height=8, results="asis"}
cat("\n\n### Fresh samples only \n\n")
plotReducedDim(sce_fresh, "UMAP", colour_by="label", text_by="label", point_size=0.25) +
  labs(title="UMAP of clusters",
       subtitle="Fresh samples only") +
  theme_ipsum_rc()

cat("\n\n### Fresh and cultured samples \n\n")
plotReducedDim(sce_all, "UMAP", colour_by="label", text_by="label", point_size=0.25) +
  labs(title="UMAP of clusters",
       subtitle="Fresh and cultured samples") +
  theme_ipsum_rc()
```

## Stratified by sample and coloured by cell cycle phase 

```{r visualise-phases-umap, fig.width=13, fig.height=8, results="asis"}
cat("\n\n### Fresh samples only {.tabset}\n\n")
plot_stratified_views(sce_fresh, "UMAP", stratify_by="Sample", colour_by="phase", text_by="label", tabset="####")

cat("\n\n### Fresh and cultured samples  {.tabset}\n\n")
plot_stratified_views(sce_all, "UMAP", stratify_by="Sample", colour_by="phase", text_by="label", tabset="####")

```


# Cluster-phase composition (stacked barplot) 

## Cell cycle by sample {.tabset}

```{r average doublet score, results="asis", fig.width=14, fig.height=10}
cat("\n\n### Fresh samples only \n\n")
as.data.frame(colData(sce_fresh)) %>%
  group_by(label, phase) %>%
  summarise(Freq=n()) %>% 
  ggplot(aes(x=label, y=Freq, fill=phase, label=Freq)) +
         geom_bar(stat="identity", position="fill") +
  labs(title="Frequency of cell phase by cluster",
       subtitle="Labels represent absolute number of cells per group",
       x="Cluster",
       y="Frequency") +
  geom_text(size=3, position = position_fill(vjust=0.5))

cat("\n\n### Fresh and cultured samples \n\n")
as.data.frame(colData(sce_all)) %>%
  group_by(label, phase) %>%
  summarise(Freq=n()) %>% 
  ggplot(aes(x=label, y=Freq, fill=phase, label=Freq)) +
         geom_bar(stat="identity", position="fill") +
  labs(title="Frequency of cell phase by cluster",
       subtitle="Labels represent absolute number of cells per group",
       x="Cluster",
       y="Frequency") +
  geom_text(size=3, position = position_fill(vjust=0.5))

```


# Save data 

```{r save data}
# Fresh samples only
file_path_fresh <- "../data/EOS_files/Fresh_only_EOS2.rds"
rownames(sce_fresh) <- rowData(sce_fresh)$Symbol
saveRDS(sce_fresh, file=file_path_fresh)
upload_file_to_iSEE(file_path_fresh)

# Frsh and cultured samples
file_path_all <- "../data/EOS_files/FvC_EOS2.rds"
rownames(sce_all) <- rowData(sce_all)$Symbol
saveRDS(sce_all, file=file_path_all)
upload_file_to_iSEE(file_path_all)
```
